---
title: "Chocolate"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, include=FALSE}
# Be sure to use your file menus at the top of R-Studio to:
# "Session -> Set working directory -> To source file location"
# Then play this chunk and the next one to get the data into R.
library(mosaic)
library(car)
library(DT)
library(pander)
library(readr)
choc <- read_csv("Data/Chocolate.csv")

choc1 <- choc %>%
mutate(Cacao = parse_number(Cocao))
choc1$Cacao <- cut(choc1$Cacao, c(42, 62, 82, 100), labels = c("Medium", "Dark", "Really Dark"))
choc1 <- select(choc1, c(Rating, Cacao, Orig, Review))

```

```{r, message = FALSE, warning = FALSE}
datatable(choc1, options = list(lengthMenu = c(3, 5, 20)))
```


----

<!-- Instructions:

 Use the Chocolate dataset and a Kruskal-Wallis Test(s) to answer an interesting question(s) that you come up with.
 
 About the dataset:
 
  * Company: Name of the company manufacturing the bar.

  * Orig: The specific geo-region of origin for the bar.

  * Review: Date of publication of the review.

  * Cocoa: Cocoa percentage (darkness) of the chocolate bar being reviewed.

  * Location: Manufacturer base country.

  * Rating: Expert rating for the bar.
      5= Elite (Transcending beyond the ordinary limits)
      4= Premium (Superior flavor development, character and style)
      3= Satisfactory(3.0) to praiseworthy(3.75) (well made with special qualities)
      2= Disappointing (Passable but contains at least one significant flaw)
      1= Unpleasant (mostly unpalatable)

   * Type: The variety (breed) of cocoa bean used, if provided.
   
   * BroadOrig: The broad geo-region of origin for the bean.

Note this dataset and description come from: © 2018 Kaggle Inc for further details visit:

[Kaggle.com](https://www.kaggle.com/rtatman/chocolate-bar-ratings/data)


If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->
**Comments to Critiquers:** 

The word "Cacao" may be spelled differently throughout this entire thing. It may even be mispelled there. I realize this and I apologize for it...


<!--End your comments to your critiquers before this line.-->

----

<!--Begin writing your Analysis after this point.-->

## Background

Dark chocolate has been one of my favorite treats for years now. Only recently did I discover that there is even darker chocolate than the chocolate that I have consumed before. I would like to see if expert ratings of chocolate match up with my taste. The dataset being used is taken from Kaggle Inc. and contains almost 11 years worth of ratings. Below you can find a link to the data as well as a short description of the relevant data. The datatable obviously doesn't contain all of the information. It only contains a fraction of what the dataset contains. If you want to know more, please visit the link posted below. 

  * Rating: Expert rating for the bar.
      
      5= Elite (Transcending beyond the ordinary limits)
      
      4= Premium (Superior flavor development, character and style)
      
      3= Satisfactory(3.0) to praiseworthy(3.75) (well made with special qualities)
      
      2= Disappointing (Passable but contains at least one significant flaw)
      
      
      1= Unpleasant (mostly unpalatable)
      
  * Cocoa: Cocoa percentage (darkness) of the chocolate bar being reviewed.
      
      For personal convenience, the percentages have been broken up into three sections: 
      1. 42-62 % is "Medium"
      2. 63-82 % is "Dark" 
      3. 83-100 % is "Really Dark"

[Kaggle.com](https://www.kaggle.com/rtatman/chocolate-bar-ratings/data)


## Hypotheses


$$
H_0: \text {All samples of ratings come from the same distribution}
$$


$$
H_a: \text {At least one distibution of ratings is stochastically different than the others}
$$

$\alpha$ = 0.05

## Test and Graphics

```{r, message = FALSE, warning = FALSE}
palette(c("tan3", "sienna" ,"saddlebrown"))
pander(kruskal.test(Rating ~ Cacao, data = choc1))
boxplot(Rating ~ Cacao, data = choc1, main = "Ratings of Chocolate according to Kaggle Inc.", ylab = "Flavor Rating of Chocolate", xlab = "Cacao concentration", col = palette()) 
legend("topright", legend = c("42-62%", "63-82%","83-100%"), bty = "n", pch = 16, col = palette())
pander(favstats(Rating ~ Cacao, data = choc1)[,-10]) 
```

## Interpretation and Conclusion

With a $p$-value that requires scientific notation to be listed, I think that is safe to conclude that the null hypothesis should be rejected at the 0.05 significance level. Therefore, there is sufficent evidence to say that one of the distributions of ratings is stochastically different than the others. Judging by the boxplot and the numerical summary, I would guess that the "Dark" category is the one that is stochastically different. The reason why I think that is because it has the highest mean and median. The "Dark" category has a mean of **3.22** and a median of **3.25**. Meaning that the chocolate of the Dark category typically has a "Satisfactory" flavor. Although, they are all pretty close to one another. Further studies could be conducted to find out once and for all which cacao percentage category is the greatest (I'm still impartial to the "Really Dark" category). In all honesty, I am a bit sad that the Really Dark didn't do as well as I would have hoped. It has an average flavor rating of **2.792** meaning that it is *almost* "Satisfactory." With these further studies, it would probably be best to have sample sizes that are remotely close to each other.   
